18 month progress report

### 2017-03-25 reports/progress_1/progress_1.tex
In progress report, complete.

Let $R_\text{b}(x,y)$, $R_\text{g}(x,y)$ and $R_\text{w}(x,y)$ be the grey value of the pixel, at position $(x,y)$, of the sample mean b/g/w images respectively. Let these images have $h$ rows and $w$ columns. The shading free b/g/w images should be completely flat, in other words all the pixels within a reference image should have the same grey value. A good guess what that shading corrected grey value should be would be the sample mean of the shading uncorrected grey values within the b/g/w images $\overline{R}_\text{b}$, $\overline{R}_\text{g}$ and $\overline{R}_\text{w}$ where
\begin{equation}
\overline{R}_j=\frac{\sum_{x=1}^w\sum_{y=1}^hR_j(x,y)}{n}
\end{equation}
for $j\in\{\text{b},\text{g},\text{w}\}$. By treating these within image sample mean grey values as the true shading free grey value, then the problem becomes a linear regression with 3 data points.

@@

To avoid dealing with dead pixels \cite{brettschneider2014spatial}, a method was considered which smoothed the reference images prior to any shading correction parameter estimation. The purpose of this was to iron out any misbehaving pixels in the reference images and to capture the shading effects parametrically. Second order panel-wise surface polynomials were fitted to the b/g/w images, this is shown in Figure \ref{fig:polynomial_shadingCorrection}. The standardised residuals from the panel-wise fit are shown in Figure \ref{fig:polynomial_residual}. A quick inspection of the residuals shows any characteristics not captured by the panel-wise fit. In the black and grey images, the residuals showed a global ring which was very unusual for the black image because it is detecting no x-rays at all. The b/g/w residuals all showed vertical stripes. The panel-wise fit can be improved by including a global polynomial fit to capture between panel effects. Higher order polynomials can be considered but overfitting may be possible as may attempt to capture all the characteristics of the b/g/w images.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_identity.png}
		\caption{Identity link, order 1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_log.png}
		\caption{Log link, order -1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_canonical_1.png}
		\caption{Inverse link, order -1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_canonical_2.png}
		\caption{Inverse link, order -2}
	\end{subfigure}
	\caption{Shading uncorrected within pixel mean and variance grey value frequency density plot of each pixel in the segmented image of the 3D printed sample. A Gamma GLM was fitted with different link functions and polynomial ordera. The solid and dotted lines are the mean response and the $\Phi(\pm 1)$ quantiles respectively.}
	\label{fig:meanVar_link}
\end{figure*}

Different shading corrections had small effects on the mean variance relationship, as shown in Figure \ref{fig:meanVar_shadingCorrection}. Due to some misbehaving pixels in the reference images, grey value outliers were possible such as negative values. Pixels with grey values less than $5.8\times 10^3$ or more than $6.6\times 10^4$ were considered outliers. Outliers were replaced by the median of its 8 nearest neighbours.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_no_shad.png}
		\caption{No shading correction}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_bw.png}
		\caption{b/w shading correction}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_bgw.png}
		\caption{b/g/w shading correction}
	\end{subfigure}
	\caption{Frequency density plot of the standardised residuals from predicting the variance given the mean in test set using a training GLM with identity link function.}
	\label{fig:glm_residual}
\end{figure*}

@@

\section{Compound Poisson}
In the mini-project, a compound Poisson model was used to model the distribution of greyvalues in a pixel \cite{ip2016inside}. The model was built by assuming that photon arrivals are a Poisson process and that each photon has random energy, distributed according to an energy spectrum. The grey value was then proportional to the sum of energies of each photon detected in some exposure time, with some Normal noise added.

@@

It is very well known that the marginal p.d.f.~$X$ is intractable. However the m.g.f.~can be written in closed form. The m.g.f.~is defined as
\begin{equation}
M(\theta)=\expectation\left[\euler^{\theta X}\right] \ .
\end{equation}
Using the property of conditional expectations
\begin{equation*}
M(\theta)=\expectation\expectation\left[\euler^{\theta X}\right|Y]
\end{equation*}
and then the properties of the Gamma distribution
\begin{equation*}
M(\theta)=\expectation\left[\left(\frac{\lambda}{\lambda-\theta}\right)^{Y\alpha}\right] \ .
\end{equation*}
Finally using the m.g.f.~of the Poisson distribution
\begin{equation}
M(\theta)=\exp\left[\nu\tau\left(\left(\frac{\lambda}{\lambda-\theta}\right)^{\alpha}-1\right)\right] \ .
\end{equation}
Moments up to a number of orders then can be obtained. For example
\begin{equation}
\expectation\left[X\right]=\frac{\alpha\nu\tau}{\lambda}
\end{equation}
\begin{equation}
\variance\left[X\right]=\frac{\alpha(\alpha+1)\nu\tau}{\lambda^2}
\end{equation}
\begin{equation}
\expectation\left[(X-\mu)^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\nu\tau}{\lambda^3}
\end{equation}
where $\expectation\left[X\right]=\mu$.

@@

For the Normal looking compound Poisson, both approximation performed well.
\begin{figure*}
\centerline{
	\centering
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_1.eps}
        \caption{Normal approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_1.eps}
        \caption{Saddlepoint approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_2.eps}
        \caption{Normal approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_2.eps}
        \caption{Saddlepoint approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_3.eps}
        \caption{Normal approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_3.eps}
        \caption{Saddlepoint approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
}
\caption{A sample of $10^4$ compound Poisson random variables was simulated. The histogram of the samples was plotted along with the density approxmation.}
\label{fig:histogram}
\end{figure*}

>>>>

Let $R_\text{b}(x,y)$, $R_\text{g}(x,y)$ and $R_\text{w}(x,y)$ be the grey value of the pixel, at position $(x,y)$, of the sample mean b/g/w images respectively. Let these images have $h$ rows and $w$ columns. The shading free b/g/w images should be completely flat, in other words all the pixels within a reference image should have the same grey value. A good guess what that shading corrected grey value should be would be the sample mean of the shading uncorrected grey values within the b/g/w images $\overline{R}_\text{b}$, $\overline{R}_\text{g}$ and $\overline{R}_\text{w}$ where
\begin{equation}
\overline{R}_j=\frac{\sum_{x=1}^w\sum_{y=1}^hR_j(x,y)}{wh}
\end{equation}
for $j\in\{\text{b},\text{g},\text{w}\}$. By treating these within image sample mean grey values as the true shading free grey value, then the problem becomes a linear regression with 3 data points.

@@

To avoid dealing with dead pixels \cite{brettschneider2014spatial}, a method was considered which smoothed the reference images prior to any shading correction parameter estimation. The purpose of this was to iron out any misbehaving pixels in the reference images and to capture the shading effects parametrically. Second order panel-wise surface polynomials were fitted to the $R_\text{b}(x,y)$, $R_\text{g}(x,y)$ and $R_\text{w}(x,y)$ images, this is shown in Figure \ref{fig:polynomial_shadingCorrection}. The standardised residuals from the panel-wise fit are shown in Figure \ref{fig:polynomial_residual}. A quick inspection of the residuals shows any characteristics not captured by the panel-wise fit. In the black and grey images, the residuals showed a global ring which was very unusual for the black image because it is detecting no x-rays at all. The b/g/w residuals all showed vertical stripes. The panel-wise fit can be improved by including a global polynomial fit to capture between panel effects. Higher order polynomials can be considered but overfitting may be possible as it may attempt to capture all the characteristics of the b/g/w images.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_identity.png}
		\caption{Identity link, order 1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_log.png}
		\caption{Log link, order -1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_canonical_1.png}
		\caption{Inverse link, order -1}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/meanVar_canonical_2.png}
		\caption{Inverse link, order -2}
	\end{subfigure}
	\caption{Shading uncorrected within pixel mean and variance grey value frequency density plot of each pixel in the segmented image of the 3D printed sample. A Gamma GLM was fitted with different link functions and polynomial orders. The solid and dotted lines are the mean response and the $\Phi(\pm 1)$ quantiles respectively.}
	\label{fig:meanVar_link}
\end{figure*}

Different shading corrections had small effects on the mean variance relationship, as shown in Figure \ref{fig:meanVar_shadingCorrection}. Due to some misbehaving pixels in the reference images, grey value outliers were possible such as negative values. Pixels with grey values less than $5.8\times 10^3$ or more than $6.6\times 10^4$ were considered outliers. Outliers were replaced by the median of its 8 post shading correction nearest neighbours.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_no_shad.png}
		\caption{No shading correction}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_bw.png}
		\caption{b/w shading correction}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width = \textwidth]{../figures/meanVar/residual_bgw.png}
		\caption{b/g/w shading correction}
	\end{subfigure}
	\caption{Frequency density plot of the standardised residuals from predicting the variance given the mean using data in a test set. The model used is a GLM with identity link function.}
	\label{fig:glm_residual}
\end{figure*}

@@

\section{Compound Poisson}
In the mini-project, a compound Poisson model was used to model the distribution of grey values in a pixel \cite{ip2016inside}. The model was built by assuming that photon arrivals are a Poisson process and that each photon has random energy, distributed according to an energy spectrum. The grey value was then proportional to the sum of energies of each photon detected in some exposure time, with some Normal noise added.

@@

It is very well known that the marginal p.d.f.~of $X$ is intractable. However the m.g.f.~can be written in closed form. The m.g.f.~is defined as
\begin{equation}
M(\theta)=\expectation\left[\euler^{\theta X}\right] \ .
\end{equation}
Using the property of conditional expectations
\begin{equation*}
M(\theta)=\expectation\expectation\left[\euler^{\theta X}\right|Y]
\end{equation*}
and then the properties of the Gamma distribution
\begin{equation*}
M(\theta)=\expectation\left[\left(\frac{\lambda}{\lambda-\theta}\right)^{Y\alpha}\right] \ .
\end{equation*}
Finally using the m.g.f.~of the Poisson distribution
\begin{equation}
M(\theta)=\exp\left[\nu\tau\left(\left(\frac{\lambda}{\lambda-\theta}\right)^{\alpha}-1\right)\right] \ .
\end{equation}
Moments up to a number of orders then can be obtained. For example
\begin{equation}
\expectation\left[X\right]=\frac{\alpha\nu\tau}{\lambda}
\end{equation}
\begin{equation}
\variance\left[X\right]=\frac{\alpha(\alpha+1)\nu\tau}{\lambda^2}
\end{equation}
\begin{equation}
\expectation\left[(X-\mu)^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\nu\tau}{\lambda^3}
\end{equation}
where $\expectation\left[X\right]=\mu$.

@@

For the Normal looking compound Poisson, both approximations performed well.
\begin{figure*}
\centerline{
	\centering
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_1.eps}
        \caption{Normal approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_1.eps}
        \caption{Saddlepoint approximation: $\nu\tau=1$, $\alpha=1$, $\lambda=1$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_2.eps}
        \caption{Normal approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_2.eps}
        \caption{Saddlepoint approximation: $\nu\tau=3$, $\alpha=400$, $\lambda=100$}
    \end{subfigure}
}
\centerline{
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/normal_3.eps}
        \caption{Normal approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=1\textwidth]{../figures/compoundPoisson/saddle_3.eps}
        \caption{Saddlepoint approximation: $\nu\tau=500$, $\alpha=80$, $\lambda=2$}
    \end{subfigure}
}
\caption{A sample of $10^4$ compound Poisson random variables was simulated. The histogram of the samples was plotted along with the density approxmation.}
\label{fig:histogram}
\end{figure*}
